---
title: "build_knowledgebase"
author: "Dongze He"
date: "5/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages

```{r}
suppressPackageStartupMessages({
  library(igraph)
  library(OmnipathR)
  library(dplyr)
  library(biomaRt)
  library(STRINGdb)
  library(ensembldb)
  library(EnsDb.Hsapiens.v86)
})
edb <- EnsDb.Hsapiens.v86
```

# load omnipath dataset

As these two databases use different annotations, I get the Entrez protein ID from `ensembldb`, if there is no corresponding entry, we have `NA`

In this step, we need to filter interactions that have no direction, or neither simulation nor inhibition

Some interactions are not included in stringDB, so we *make* the combined score according to #references, #sourcese and curation effort.

```{r}
# get interactions from all databases
omni = import_omnipath_interactions(
    organism = 9606,
    datasets = c('omnipath', 'pathwayextra', 'kinaseextra', 'ligrecextra', 'dorothea','tf_target', 'mirnatarget', 'tf_mirna', 'lncrna_mrna')
)

# This is the default from the database location https://omnipathdb.org/interactions
omni_human = import_omnipath_interactions(
    organism = 9606,
)

table((omni$consensus_stimulation + omni$consensus_inhibition))


# keep only interactions with clear direction
omni = omni[(omni$consensus_stimulation + omni$consensus_inhibition) == 1,]

# some entries is out of stringdb, so we make the score 
table(omni$n_resources)
table(omni$curation_effort)
  
#june 2021
#omni$combined_score = (omni$n_references/max(omni$n_references) + omni$n_resources/max(omni$n_resources) + omni$curation_effort/max(omni$curation_effort))/3

#Oct 2021 (stopped working )
# The combined score is the two values given by the database ... (log score to only penalize low # of citations. But anything high should be reliable)
omni$combined_score = ( log(1+omni$n_resources)/log(max(omni$n_resources)) + log(1+omni$curation_effort)/log(max(omni$curation_effort)))/2

# 0.15 is the minimum score in stirngDB
omni$combined_score[omni$combined_score < 0.15] = 0.15

# get ensembl ID from ensembl db
uniprot_to_ensembl = proteins(EnsDb.Hsapiens.v86, filter = UniprotFilter(c(omni$source, omni$target)))
rownames(uniprot_to_ensembl) = uniprot_to_ensembl$uniprot_id
omni$source_ensembl = uniprot_to_ensembl[match(omni$source, uniprot_to_ensembl$uniprot_id),]$protein_id 
omni$target_ensembl = uniprot_to_ensembl[match(omni$target, uniprot_to_ensembl$uniprot_id),]$protein_id 

# as we need to match the two datasets using two columns, source and target, so we combined the to into one
omni$st = paste0(omni$source_ensembl, omni$target_ensembl)

```

# STRINGdb

we read in STRINGdb and process the ID. As STRINGdb is undirected, we swap the from and two and rbind the interactions with its swapped one.

```{r}
# instantiate stringdb
string_db <- STRINGdb$new( version="11.5", species=9606,
 score_threshold=0, input_directory="")

# get interactions
string_db_interactions = string_db$get_interactions(string_db$get_aliases()$STRING_id)

# remove the prefix in the entrez ID
string_db_interactions$from <- sub("9606.", "", string_db_interactions$from)
string_db_interactions$to <-sub("9606.", "", string_db_interactions$to)

# swap from and to, then rbind them
string_db_interactions_reverse = string_db_interactions[, c(2,1,3)]
colnames(string_db_interactions_reverse) = colnames(string_db_interactions)
string_db_interactions = rbind(string_db_interactions, string_db_interactions_reverse)

# match omnipath using source and target ID (the combined one)
string_db_interactions$st = paste0(string_db_interactions$from, string_db_interactions$to)
string_db_interactions = string_db_interactions[which(string_db_interactions$st %in% omni$st),]
rm(string_db, string_db_interactions_reverse)

# scale combined score, and aggregate the scores if the source and target names are same
string_db_interactions$combined_score = string_db_interactions$combined_score/1000
string_db_interactions = aggregate(combined_score ~ from + to, string_db_interactions, mean)

# set rownames for the further steps
rownames(string_db_interactions) = paste0(string_db_interactions$from, string_db_interactions$to)
```

# Combine two datasets

```{r}
# using source and target IDs, we find the matched rows and append the combined score
omni$combined_score[omni$st %in% rownames(string_db_interactions)] = string_db_interactions[omni$st[omni$st %in% rownames(string_db_interactions)],]$combined_score

# we define sign: 1 is stimulated, -1 is inhibition
omni$sign = omni$consensus_stimulation - omni$consensus_inhibition
```

# Take one interested columns as our PPI

```{r}
PPI = omni[, c("source_genesymbol", "target_genesymbol", "sign", "combined_score")]
colnames(PPI) <- c("source_name", "target_name", "sign", "weight")
saveRDS(PPI, "PPI_graph.rds")
```
